---
title: "Practical Machine Learning - course project"
author: "Kleber"
date: '2022-10-26'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the peer review part of the project for the course "Practical Machine Learning". The goal was to make a model 
to predict the variable "classe" using the data provided by the course.

## Analysis

These are the steps that i have done to reach a model that i could trust:

1. I loaded the training set and removed the column "user_names", which is useless;
2. I replaced the strings "#DIV/0!" and "<NA>" for NA;
3. I saw the data set had a lot of missing values (NA). I tried to estimate a tree model but the function said that
was not possible because the missing data. So, a removed all the variables which have some missing data. I used in the
model just 56 predictors because of that;
4. First i estimated the simple tree model using the method "rpart" of the "train" function (package "caret"). I did the
confusion matrix for the training data and the accuracy was 0.5332. Then, i estimated a new model using bagging, by the 
method "treebag" and the accuracy of this model was 0.9988. I chosen the last model;
5. Finally, i predict the "classe" variable using the testing data set and the tree bag model.


```{r codification, echo=T}
# load libraries
library(caret)
library(readr)
#_______________________________________________________________________________________________________________________
# load the training dataset
#_______________________________________________________________________________________________________________________
# load training set
pml_training <- read_csv("pml-training.csv", 
                         col_types = cols(...1 = col_skip()))
# convert to data.frame
pml_training <- as.data.frame(pml_training)
# remove the column "user_name"
pml_training <- pml_training[,-1]
# identify the columns with any NA
nomes_col <- names(apply(pml_training, 2, function (x) any(is.na(x)))[apply(pml_training, 2, 
                                                                            function (x) any(is.na(x)))==TRUE])
# remove coluns with NA
pml_training_2 <- pml_training[,which(!colnames(pml_training)%in%nomes_col)]
# replace "#DIV/0!" by NA
pml_training_2[pml_training_2=="#DIV/0!"] <- NA
# replace <NA> by NA
pml_training_2[pml_training_2=="<NA>"] <- NA
# remove the column 'cvtd_timestamp' because is causing error in the prediction funcion
pml_training_2[,c("cvtd_timestamp","new_window")] <- NULL
#_______________________________________________________________________________________________________________________
# load the testing dataset
#_______________________________________________________________________________________________________________________
# load testing set
pml_testing <- read_csv("pml-testing.csv", 
                         col_types = cols(...1 = col_skip()))
# convert to data.frame
pml_testing <- as.data.frame(pml_testing)
# remove the column "user_name"
pml_testing <- pml_testing[,-1]
# identify the columns with any NA
nomes_col <- names(apply(pml_testing, 2, function (x) any(is.na(x)))[apply(pml_testing, 2, 
                                                                            function (x) any(is.na(x)))==TRUE])
# remove coluns with NA
pml_testing_2 <- pml_testing[,which(!colnames(pml_testing)%in%nomes_col)]
# replace "#DIV/0!" by NA
pml_testing_2[pml_testing_2=="#DIV/0!"] <- NA
# replace <NA> by NA
pml_testing_2[pml_testing_2=="<NA>"] <- NA
# remove the column 'cvtd_timestamp' because is causing error in the prediction funcion
pml_testing_2[,c("cvtd_timestamp","new_window")] <- NULL
#_______________________________________________________________________________________________________________________
# training the model
#_______________________________________________________________________________________________________________________
# using 10 k-fold
train_control<- trainControl(method="cv", number=10, savePredictions = TRUE)
# training the classic tree model
modFit_1 <- train(classe~., method="rpart", data=pml_training_2, trControl=train_control)
# confusion matrix of the training set prediction
confusionMatrix(modFit_1)
# predict the model 1 with the test set
pred_1 <- predict(modFit_1, data=pml_testing_2)
# training the bag of trees
modFit_2 <- train(classe~., method="treebag", data=pml_training_2, trControl=train_control)
# confusion matrix of the training set prediction
confusionMatrix(modFit_2)
# predict the model 2
pred_2 <- predict(modFit_2, data=pml_testing_2)
```
